---
title: "Excercise 1"
author: "Dennis Jakob"
date: "11/4/2019"
output: html_document
---

# Task 1
Create a vector of *100 numbers* from a normal distribution with $\mu$ = 0 and $\sigma ^2$ = 1. (see ?rnorm). Store this in an object called “myNorm.01.100”.

```{r}
set.seed(10)

myNorm.01.100 <- rnorm(
  n = 100,  
  mean = 0, 
  sd = 1     
)
```

# Task 2
Calculate the different measures of location and spread. (see?mean).

```{r}
# calculate mean
myMean <- mean(myNorm.01.100)
# calculate median
myMedian <- median(myNorm.01.100)
# create mode function
MyMode <- function (x) {  
  x.round <- round(x,1)  
  x.table <- table(x.round) 
  mode <- x.table[x.table == max(x.table)] 
  return(mode)
} 
# calculate mode
myMode <- MyMode(myNorm.01.100)
# calculate variance
myVar <- var(myNorm.01.100)
# calculate standard deviation
mySD <- sd(myNorm.01.100)
# calculate range
myRange <- range(myNorm.01.100)
```

# Task 3
Plot the 100 observations from tow normal distributions, the first one with $\mu = 5$ and $\sigma ^2 = 3$ and the second with $\mu = 5$ and $\sigma ^2 = 10$ (“myNorm.53.100” and “myNorm.510.100”). What can you say about the difference between the two? (See ?hist)

```{r}
set.seed(10)
myNorm.53.100 <- rnorm(100, 5, 3)

set.seed(10)
myNorm.510.100 <- rnorm(100, 5, 10)

par(mfrow = c(1,2)) 
hist(myNorm.53.100)
hist(myNorm.510.100)
```

# Task 4
Calculate and plot the z-scores for both normal distributions. Explain your results.

```{r}
# calculate z-score for myNorm.53.100
myNorm.53.100.zscore <- (myNorm.53.100 - mean(myNorm.53.100)) / sd(myNorm.53.100)
scale(myNorm.53.100)
# calculate z-score for myNorm.510.100
myNorm.510.100.zscore <- (myNorm.510.100 - mean(myNorm.510.100)) / sd(myNorm.510.100)
scale(myNorm.510.100)

par(mfrow = c(1,2)) 
plot(scale(myNorm.53.100))
plot(scale(myNorm.510.100))
```

# Task 5
Load the table "bioenv-1.csv" and histogram columns a, b and c. (see ?read.csv).
1. Make sure R understands that your data are numeric. (see?class)
2. Calculate the various measures of location spread discussed in the lecture. What can you say about these distributions?
3. Transform the non-normal distributions such that they approach normality. Which transformations did you use and why?

```{r}
# 1.
bioenv1 <- read.csv(
  "statsI/bioenv-1.csv",
  header = T,
  sep = "\t", 
  row.names = 1)
bioenv1 <- bioenv1[bioenv1$a != "no data", ] # subsets dataframe by deleting lines with missing data
bioenv1$a <- as.integer(bioenv1$a)
```

```{r}
# 2.
bioenv1.analysis <- data.frame()
for (i in 1:ncol(bioenv1)) {
  bioenv1.analysis[1,i] <- mean(bioenv1[,i])
  bioenv1.analysis[2,i] <- median(bioenv1[,i])
  bioenv1.analysis[3,i] <- MyMode(bioenv1[,i])
  bioenv1.analysis[4,i] <- var(bioenv1[,i])
  bioenv1.analysis[5,i] <- sd(bioenv1[,i])
}
rownames(bioenv1.analysis) <- (c("mean", "median", "mode", "var", "SD"))
colnames(bioenv1.analysis) <- colnames(bioenv1)

bioenv1.range <- list()
for (i in 1:ncol(bioenv1)) {
  bioenv1.range[[i]] <- range(bioenv1[,i])
}
```

```{r}
# 3.

par(mfrow = c(1,3))
hist(bioenv1$a)
hist(sqrt(bioenv1$a))
hist(log(bioenv1$a)) # brings it closer to normal distribution

par(mfrow = c(1,3))
qqnorm(bioenv1$a)
qqline(bioenv1$a)
qqnorm(sqrt(bioenv1$a), main = "sqrt transform")
qqline(sqrt(bioenv1$a), main = "log transform")
qqnorm(log(bioenv1$a + 0.1))
qqline(log(bioenv1$a + 0.1))

# species "b"
par(mfrow = c(1,3))
hist(bioenv1$b)
hist(sqrt(bioenv1$b))
hist(log(bioenv1$b)) # brings it closer to normal distribution

par(mfrow = c(1,3))
qqnorm(bioenv1$b)
qqline(bioenv1$b)
qqnorm(sqrt(bioenv1$b), main = "sqrt transform")
qqline(sqrt(bioenv1$b))
qqnorm(log(bioenv1$b + 0.1), main = "log transform")
qqline(log(bioenv1$b + 0.1))

# species "c"
par(mfrow = c(1,3))
hist(bioenv1$c)
hist(sqrt(bioenv1$c))
hist(log(bioenv1$c)) # brings it closer to normal distribution

par(mfrow = c(1,3))
qqnorm(bioenv1$c)
qqline(bioenv1$c)
qqnorm(sqrt(bioenv1$c), main = "sqrt transform")
qqline(sqrt(bioenv1$c))
qqnorm(log(bioenv1$c + 0.1), main = "log transform")
qqline(log(bioenv1$c + 0.1))
```

# Task 6
Calculate the coefficients of variation for your distributions. Can you compare this statistic across samples with different locations and spreads? Why?

```{r}
for (i in 1:ncol(bioenv1)) {
  bioenv1.analysis[6,i] <- sd(bioenv1[,i])/mean(bioenv1[,i])
}
rownames(bioenv1.analysis) <- paste(c(rownames(bioenv1.analysis[1:(length(rownames(bioenv1.analysis))-1),]), "CV"), sep = "")
```

# Task 7
Find out how to calculate the geometric mean in R. Do so for your various distributions.

```{r}
for (i in 1:ncol(bioenv1)) {
  bioenv1.analysis[7,i] <- prod(bioenv1[,i] + 1 )^(1/length(bioenv1[,i])) -1
}
rownames(bioenv1.analysis) <- paste(c(rownames(bioenv1.analysis[1:(length(rownames(bioenv1.analysis))-1),]), "g.mean"), sep = "")
```



