---
title: "Excercise 2"
author: "Dennis Jakob"
date: "11/5/2019"
output: html_document
---

# Task 1
Plot a normal distribution N(0, 1) and the t-student distributions with 2, 5, 30 and 100 degrees of freedom. What happens when n increases? What parameter of the t distribution does this affect? (see ?seq, ?dnorm and ?plot)

```{r}
# plot the density function of N(0,1)
xseq <- seq(-5,5,0.01)
xdnorm.0.1 <- dnorm(x=xseq,
                    mean=0,
                    sd=1) 
xdt.2 <- dt(xseq,df=2)
xdt.5 <- dt(xseq,df=5)
xdt.30 <- dt(xseq,df=30)
xdt.100 <- dt(xseq,df=100)

# plot the normal distribution
plot(xseq,
     xdnorm.0.1,
     type="l",
     lwd=2.5)
# add the density function of T(df), f = 2,5,30,100 in the plot
points(xseq,xdt.2,col="blue", type="l")
points(xseq,xdt.5,col="red", type="l")
points(xseq,xdt.30,col="green", type="l")
points(xseq,xdt.100,col="orange", type="l")
```
The higher the degrees of freedom get, the more does the Student distribution approach a nomal distribution. $\rightarrow$ the number of replicates of trials is increased
The the probability for the mean increases, the tails get narrower.

# Task 2
Plot two binomial distributions: B(20,0.5) and B(50,0.3). Can you approximate any of these with a normal distribution? (see ?rbinom)

```{r}
p = 0.5 # probability of success
q = 1-p # complement of success
n = 1000 # number "experiments"

# number of trials = 20
set.seed(3)
binom.1K.20.5 <- rbinom(n=n,
                        size=20,
                        prob=p)
hist(binom.1K.20.5,
     freq=T)

# number of trials = 50
set.seed(3)
binom.1K.50.5 <- rbinom(n=n,
                        size=50,
                        prob=p)
hist(binom.1K.50.5,
     freq=T)

# binom.1K.50.5 will be used for the approximation
binmo.1K.50.5.mean <- 50*p
binmo.1K.50.5.sd <- sqrt(50*p*q)

xseq <- seq(0,50,0.1)
norm.approx <- dnorm(x=xseq,
                     mean=binmo.1K.50.5.mean,
                     sd=binmo.1K.50.5.sd)

set.seed(3)
binom.1K.50.5 <- rbinom(n=n,
                        size=50,
                        prob=p)
hist(binom.1K.50.5,
     freq=F)
lines(xseq,
      norm.approx,
      col="red")
```

# Task 3
Plot a Poisson distribution Pois(12). What does the value of lambda control. How could you approximate it to a normal distribution? (see ?rpois)

```{r}
# Overlay histograms
hist(rpois(1000,lambda=12),xlim=c(0,30),ylim=c(0,250),col=rgb(0,0,0.5,0.5))
hist(rnorm(1000,mean=12,sd=sqrt(12)),add=T,col=rgb(0.5,0,0,0.5))
```
Lambda describes the means of the distribution. Means and variance are equal.

# Task 4
Given X ~ N(0,1), what is the probability of obtaining x from X, such that x > 1.96. Compare this with T(i) for i = 10, 20, 30, 100 and 100000. (see ?pnorm and ?pt)

```{r}
plot
pnorm(q=1.96,mean=0,sd=1,lower.tail=F) # 0.0249979
pt(q=1.96,df=10) # 0.9607819
pt(q=1.96,df=20) # 0.9679609
pt(q=1.96,df=30) # 0.9703288
pt(q=1.96,df=100) # 0.9736105
pt(q=1.96,df=100000) # 0.9750007
```
The probability is lowered in the t-distribution with df = 10.
For a high df the t-distribution approaches the normal distribution.

# Task 5
From table bioenv-2, find the covariance and correlation between the species abundance variables and temperature. Are any of these correlations significant? Do this with parametric (pearson) and nonparametric (spearman) methods.
Follow up by plotting these variables against one another (see ?pairs). What do you observe?

```{r}
bioenv2 <- read.csv(
  "statsI/bioenv-2.csv",
  header = T,
  sep = ",", 
  row.names = 1)
head(bioenv2)
str(bioenv2)

# species vs temperature
bioenv2.analysis <- data.frame()
for (i in 1:(ncol(bioenv2)-1)) {
  bioenv2.analysis[1,i] <- cov(x = bioenv2[,i],
                               y = bioenv2$Temperature,
                               method = "pearson")
  bioenv2.analysis[2,i] <- cov(x = bioenv2[,i],
                               y = bioenv2$Temperature,
                               method = "spearman")
  bioenv2.analysis[3,i] <- cor(x = bioenv2[,i],
                               y = bioenv2$Temperature,
                               method = "pearson")
  bioenv2.analysis[4,i] <- cor(x = bioenv2[,i],
                               y = bioenv2$Temperature,
                               method = "spearman")
}
colnames(bioenv2.analysis) <- colnames(bioenv2[,1:4])
rownames(bioenv2.analysis) <- c("cov.pearson", "cov.spearman", "cor.pearson", "cor.spearman")

# correlation test
bioenv2.cor.test.pearson <- list()
bioenv2.cor.test.spearman <- list()
for (i in 1:(ncol(bioenv2)-1)) {
  bioenv2.cor.test.pearson[[i]] <- cor.test(x=bioenv2[,i],
                                            y=bioenv2$Temperature,
                                            method="pearson")
  bioenv2.cor.test.spearman[[i]] <- cor.test(x=bioenv2[,i],
                                             y=bioenv2$Temperature,
                                             method="spearman")
}

# plotting variables against each other
par(mfrow = c(2,2))
for (i in 1:(ncol(bioenv2)-1)) {
  plot(bioenv2$Temperature,bioenv2[,i])
}
par(mfrow = c(1,1))
pairs(bioenv2)
```

# Task 6
Calculate and plot a linear regression for an appropriate pair of variables in your data table. How much are the explained and residual variability? Are the residuals normally distributed? Perform residual plots.

```{r}
bioenv2.lm <- list()
par(mfrow=c(2,2))
for (i in 1:(ncol(bioenv2)-1)) {
  bioenv2.lm[[i]] <- lm(bioenv2[,i]~bioenv2$Temperature)
  plot(bioenv2$Temperature,bioenv2[,i])
  abline(bioenv2.lm[[i]],lwd=2,col="red")
}
# summary(bioenv2.lm[[1]])

# regression is the explained part
SSreg <- var(bioenv2.lm[[4]]$fitted.values) 
SStotal <- var(bioenv2$d)
R2 <- SSreg/SStotal
# everything that is not explained is residual
residual <- SStotal - SSreg
```

# Task 7
Calculate the total, residual and fitted variability.

```{r}
bioenv2.variability <- data.frame()
for (i in 1:(ncol(bioenv2)-1)) {
  bioenv2.variability[1,i] <- var(residuals(bioenv2.lm[[i]]))
  bioenv2.variability[2,i] <- var(fitted(bioenv2.lm[[i]]))
  bioenv2.variability[3,i] <- var(bioenv2[,i])
  bioenv2.variability[4,i] <- bioenv2.variability[1,i] + bioenv2.variability[2,i]
  bioenv2.variability[5,i] <- bioenv2.variability[2,i]/(bioenv2.variability[1,i] + bioenv2.variability[2,i])
}
colnames(bioenv2.variability) <- colnames(bioenv2[,1:4])
rownames(bioenv2.variability) <- c("r.variability", "f.variability", "t.variability", "r+f.variability", "f/r+f.variability")
```









