---
title: "Exercise 3"
author: "Dennis Jakob"
date: "11/6/2019"
output: html_document
---

# to do

- data.table
- ggplot2
- tidyverse

# Task 1
Load table bioenv3 and perform the following hypothesis tests:
1) Can we claim that the population mean of species “a” is bigger than 10?
2) Is the population mean of species “b” different from 10?
3) Are the population means of species “a” and “b” different from each other? What do you see in the degrees of freedom? Try setting the var.equal argument to TRUE. Use ?t.test to find out what has happened.

```{r}
bioenv3 <- read.csv(
  "statsI/bioenv-3.csv",
  header = T,
  sep = ",", 
  row.names = 1)
head(bioenv3)
str(bioenv3)

bioenv3.analysis <- data.frame()
for (i in 1:2) {
  # calculate mean and sd
  bioenv3.analysis["mean",i] <- mean(bioenv3[,i])
  bioenv3.analysis["SD",i] <- sd(bioenv3[,i])
  # manually compute the t statistic and the p-value 
  bioenv3.analysis["t.stat",i] <- (bioenv3.analysis[1,i]-10 ) / (bioenv3.analysis[2,i] / sqrt(length(bioenv3[,i])))
  bioenv3.analysis["p.t",i] <- pt(bioenv3.analysis[3,i], df = length(bioenv3[,i])-1, lower.tail = F)
  bioenv3.analysis["p.two.t",i] <- bioenv3.analysis[4,i]*2
}
colnames(bioenv3.analysis) <- colnames(bioenv3[,1:2])

t.stat.a.b <- (bioenv3.analysis[1,1] - bioenv3.analysis[1,2])/sqrt((bioenv3.analysis[2,1]^2)/length(bioenv3$a) + (bioenv3.analysis[2,2]^2)/length(bioenv3$b))
p.t.stat.a.b <- pt(t.stat.a.b,df=length(bioenv3$a)+length(bioenv3$b),lower.tail=T)

# visualize the test
par(mfrow = c(1,2))
for (i in 1:2) {
  plot(seq(-4, 4 , by = 0.1),
       dt(seq(-4, 4, by = 0.1), df=length(bioenv3[,i])-1),
       type = "l")
  abline(v = bioenv3.analysis[3,i])
}

# do the test using a existing function from R 
bioenv3.t.test <- list()
bioenv3.t.test[[1]] <- t.test(x= bioenv3$a,
                              mu = 10,
                              alternative = "greater")
bioenv3.t.test[[2]] <- t.test(x= bioenv3$b,
                              mu = 10,
                              alternative = "greater")
bioenv3.t.test[[3]] <- t.test(x= bioenv3$a,
                              mu = 10,
                              alternative = "two.sided")
bioenv3.t.test[[4]] <- t.test(x= bioenv3$b,
                              mu = 10,
                              alternative = "two.sided")
bioenv3.t.test[[5]] <- t.test(bioenv3$a,
                              bioenv3$b,
                              alternative="less")
bioenv3.t.test[[6]] <- t.test(bioenv3$a,
                              bioenv3$b,
                              alternative="less",
                              var.equal=T)
```

1) No, we cannot claim that the population mean of species "a" is higher than 10. p = 0.15 > $\alpha=0.05$. $H_{0}$ can not be rejected.
2) It is likely that the population mean of species "b" is different from 10. p = 2.97e-10 < $\alpha=0.05$. $H_{0}$ can be rejected. 
3) The population means of species “a” and “b” are different from each other. The degrees of freedom is reduces with var.equal=F because is uses dfs for the calculation of the standard deviation.

# Task 2
Estimate the statistical power of the previous tests.
statistical power: the probability of accepting the alternative hypothesis (H1) when it is true.

```{r}
require('pwr')
bioenv3.pwr <- list()
d.1 = sqrt((bioenv3.analysis[1,1]-10)^2)/sqrt(bioenv3.analysis[2,1]^2)
bioenv3.pwr[[1]] <- pwr.t.test(n=length(bioenv3$a), d=d.1, sig.level=0.05, power=NULL, type="one.sample", alternative="greater")
bioenv3.pwr[[2]] <- pwr.t.test(n=length(bioenv3$a), d=d.1, sig.level=0.05, power=NULL, type="one.sample", alternative="two.sided")
d.3 = sqrt((bioenv3.analysis[1,1]-bioenv3.analysis[1,2])^2)/sqrt((bioenv3.analysis[2,1]^2+bioenv3.analysis[2,2]^2/2))
bioenv3.pwr[[3]] <- pwr.t.test(n=length(bioenv3$a)+length(bioenv3$ab), d=d.3, sig.level=0.05, power=NULL, type="two.sample", alternative="less")
bioenv3.pwr[[1]]$power
bioenv3.pwr[[2]]$power
bioenv3.pwr[[3]]$power
```

# Task 3
Generate a dataset of size n=35 from a normal distribution N(3,5). Estimate the mean and its confidence interval for a 95% confidence level.

```{r}
# create data set
set.seed(1)
mynorm.35.3.5 <- rnorm(35,3,5)

mynorm.analysis <- data.frame()
mynorm.analysis["mean",1] <- mean(mynorm.35.3.5)
mynorm.analysis["sd",1] <- sd(mynorm.35.3.5)
mynorm.analysis["s.error",1] = qnorm(0.95, mean=0, sd=1)*(mynorm.analysis["sd",]/sqrt(35))
mynorm.analysis["up.lim",1] <- mynorm.analysis["mean",] + mynorm.analysis["s.error",]
mynorm.analysis["low.lim",1] <- mynorm.analysis["mean",] - mynorm.analysis["s.error",]

# double check
x <- seq(0,6,by=0.01)
plot(x,
     dnorm(x,
           mynorm.analysis["mean",],
           mynorm.analysis["sd",]/sqrt(35)),type="l")
abline(v=mynorm.analysis["up.lim",])
abline(v=mynorm.analysis["low.lim",])

pnorm(mynorm.analysis["up.lim",],
      mynorm.analysis["mean",],
      mynorm.analysis["sd",]/sqrt(35))
pnorm(mynorm.analysis["low.lim",],
      mynorm.analysis["mean",],
      mynorm.analysis["sd",]/sqrt(35))
```

# Task 4
Get a sample of size n=50 for the following 3 distributions: N(4,5), N(6,5) and N(12,4.5). Test whether the means of these samples have a significant difference.

```{r}
set.seed(1)
mydata <- list()
mydata[["a"]] <- rnorm(50,4,5)
mydata[["b"]] <- rnorm(50,6,5)
mydata[["c"]] <- rnorm(50,12,4.5)

mydata[["groups"]] = factor(rep(c("a","b","c"), each = 50))
mydata[["mydata"]] <- c(mydata[["a"]], mydata[["b"]], mydata[["c"]])

mydata.df <- data.frame(mydata[["mydata"]],mydata[["groups"]])

plot(mydata[["mydata"]] ~ mydata[["groups"]])

# boxplot example
par(mfrow=c(2,1),
    mar=c(1,1,1,1))
boxplot(mydata[["a"]], horizontal=T)
hist(mydata[["a"]])

# test homogeneity of variance (assumption of the ANOVA)
mydata[["bartlett"]] <- bartlett.test(mydata[["mydata"]], mydata[["groups"]]) # the null hypothesis is that the variances are homogeneous

# ANOVA
mydata[["anova"]] <- aov(mydata[["mydata"]] ~ mydata[["groups"]], data=mydata.df)
summary(mydata[["anova"]])

mydata[["pairwise.t"]] <- pairwise.t.test(mydata[["mydata"]], mydata[["groups"]], p.adjust="bonferroni")

# Other option is applying the function TukeyHSD().
mydata[["tukey"]] <- TukeyHSD(mydata[["anova"]], conf.level=0.95)
```

# Task 5
Using the data from table bioenv3, test if there is an association between the abundance of species “a” and the different categorical variables.

```{r}
bioenv3 <- read.csv(
  "statsI/bioenv-3.csv",
  header = T,
  sep = ",", 
  row.names = 1)
head(bioenv3)
str(bioenv3)

# chi squared and fisher's exact tests
bioenv3$Biofilm[bioenv3$Biofilm=="Present"]="present"
bioenv3 <- droplevels(bioenv3)

bioenv3.test <- list()
bioenv3.test[["chisq"]] <- chisq.test(table(bioenv3$Salinity,bioenv3$Biofilm))
bioenv3.test[["fisher"]] <- fisher.test(table(bioenv3$Salinity,bioenv3$Biofilm))
```

# Task 6
Generate a sample of size n=50 from two Poisson distributions: Pois(3) and Pois(7). Test by randomization if the means of these two distributions differ.

```{r}
# Test by randomization if the means of two samples differ. 
# First, we generate two samples of size n = 50 from two different Poisson distributions (Pois(3) and Pois(5)) and store their elements in one vector.

pois <- list()
pois[["rpois.50.3"]] <- rpois(50,3)
pois[["rpois.50.5"]] <- rpois(50,5)
pois[["p.comb"]] <- c(pois[["rpois.50.3"]], pois[["rpois.50.5"]])

# Our null hypothesis is that there is no difference between the means.
# Let's compute the null hypothesis distribution: 

rand <- vector(mode="numeric",length=10000)

for (i in 1:length(rand)) {                          # Here we are reordering the vector.
  tmp <- sample(pois[["p.comb"]])                    # Now the elements of both samples will have random positions in the vector.
  rand[i] <- mean(tmp[1:50]) - mean(tmp[51:100])     # After the samples are mixed we compute the difference of the means.
}                                                    # By repeating this several times, we can get an estimation of the null hypothesis distribution.

# Plot the null hypothesis distribution
par(mfrow=c(1,1))
hist(rand)

r.stat <- mean(pois[["p.comb"]][1:50]) - mean(pois[["p.comb"]][51:100])
r.stat
abline(v=r.stat,col="red")

# Let's see how the distribution of H0 and H1 look like 
mean_diff <- vector (mode="numeric",length=10000)

for ( i in 1:length(mean_diff)) {
  set.seed(i)
  pois[["p.comb.r"]] <- c(rpois(50,3),rpois(50,5))
  mean_diff[i] <- mean(pois[["p.comb.r"]][1:50]) - mean(pois[["p.comb.r"]][51:100])
}

hist(mean_diff)

hist(rand,
     col=rgb(0,0,0.5,0.5), 
     main="H0 and H1 distributions",
     xlab = "mean difference",
     xlim=c(min(mean_diff),max(rand))
)

hist(mean_diff, 
     add=T,
     col=rgb(0.5,0,0,0.5)
)

# See where the real difference falls under the null hypothesis
r.stat <- mean(pois[["p.comb.r"]][1:50]) - mean(pois[["p.comb.r"]][51:100])
r.stat
abline(v=r.stat,col="red")
```


