---
title: "Ex4"
author: "Dennis Jakob"
date: "1/13/2020"
output: html_document
---
```{r}
require(vegan)
require(corrgram)
require(viridis)
data("varespec")
```


1) Analyze the distribution and correlation of the variables of the varespec dataset 
1.1) What is in varespec?
```{r}
# inspect data
?varespec
dim(varespec)
str(varespec)

# check data wirh correlelogram
corrgram(varespec,
         lower.panel = panel.shade,
         upper.panel = NULL,
         text.panel = panel.txt,
         col.regions = viridis)
```

2) Perform a PCA with this data
```{r}
# base R
varspec.pca <- prcomp(varespec,
                      scale. = T)
biplot(varspec.pca)

# vegan rda (redundancy analysis)
varespec.rda <- rda(varespec)
varespec.rda.scale <- rda(varespec,
                          scale = T)
```

3) Plot the objects and variables in the first and second PCs, and in second and third PC 
3.1) What information is inside your PCA object?
```{r}
# PC1 and PC2
biplot(varespec.rda,
       choices = c(1,2),
       scaling = "species",
       display = c("sites", "species"))

# PC2 and PC3
biplot(varespec.rda,
       choices = c(2,3),
       scaling = "species",
       display = c("sites", "species"))

# PC1 and PC2 scaled
biplot(varespec.rda.scale,
       choices = c(1,2),
       scaling = "species",
       display = c("sites", "species"))

# PC2 and PC3 scaled
biplot(varespec.rda.scale,
       choices = c(2,3),
       scaling = "species",
       display = c("sites", "species"))

```


4) How much variability is explained by the first three PCs?
```{r}
summary(varespec.rda)
summary(varespec.rda.scale)
```

5) What is the difference between standardizing and not standardizing the data matrix?

-
-

6) Which are the variables that contribute the most to the first PC?

- Pleuschr
- Cladstel

7) After plotting the first 2 PCs what can you say about sampling site 28?
7.1) And what about the correlation between variables? (consider type 1 and 2 scaling).

- Pleuschr is highest at site 28
- you cannot really see it with type 2 scaling
