---
title: "Ex_CCA"
author: "Dennis Jakob"
date: "1/15/2020"
output: html_document
---
# CCA exercise
1) What is the difference between the CCA and RDA?
- RDA
-- used for data with linear relations
-- it uses the origin data
-- shows variation

- CCA
-- unimodal/non-linear data can be used by performing Chi-squared analysis
-- uses the Chi-squared
-- shows inertia

2) Perform a CCA on the varespec and varchem datasets (response and explanatory respectively).
```{r}
require(vegan)
require(corrgram)
data("varespec")
data("varechem")
?varespec
?varechem

# perform and plot CCA varespec ~ varechem
varespec.cca <- cca(formula = varespec ~ .,
             data = varechem)
```

3) Make a triplot (objects, and response and explanatory variables) for the first two canonical axes.
 3.1) How much inertia is captured by these axes?
```{r}
ordiplot(varespec.cca,
         type="t",
         scaling=1) # Type 1 scaling emphasises the relationships among objects
ordiplot(varespec.cca,
         type="t",
         scaling=2) # Type 2 scaling emphasises the relationships among response variables
ordiplot(varespec.cca,
         choices = c(1,2),
         type="t") # scaling 2 is default

varespec.cca$CCA$tot.chi
(sum(varespec.cca$CCA$eig[1:2])/sum(varespec.cca$CCA$eig))*100
```

4) What are the values of the constrained and total inertia? What does this mean? What is inertia in this case (compare it to the inertia reported in your RDA object).

```{r}
varespec.cca
# total inertia
varespec.cca$tot.chi
# unconstrained inertia
varespec.cca$CA$tot.chi
# constrained inertia
varespec.cca$CCA$tot.chi
# proportion of constrained inertia (inertia means variance)
varespec.cca$CCA$tot.chi/varespec.cca$tot.chi

varespec.rda <- rda(formula = varespec ~ .,
                    data = varechem)
varespec.rda
```

5) What is the correlation between the first canonical axes and the explanatory matrix?
```{r}
# inertia captured in the CCA axes
screeplot(varespec.cca)
sum(varespec.cca$CCA$eig[1:2])/sum(varespec.cca$CCA$eig)
```


6) Test the significance of the overall solution and for the first two axes.
```{r}
set.seed(15081992)
# test significance of the overall solution 
anova.cca(varespec.cca)

# test significance of the association between each canonical axes and explanatory variables
varespec.cca.anova.axis <- anova.cca(varespec.cca,
                                     by="axis")
varespec.cca.anova.axis

CCA1 <- varespec.cca$CCA$u[,"CCA1"]
corrgram(cbind(CCA1, varechem), lower.panel = panel.shade, upper.panel = NULL, text.panel = panel.txt)
CCA2 <-varespec.cca$CCA$u[,"CCA2"]
corrgram(cbind(CCA2, varechem), lower.panel = panel.shade, upper.panel = NULL, text.panel = panel.txt)
# analyze multi linear correlation coefficients
varespec.cca.mlm <- as.mlm(varespec.cca)
varespec.cca.mlm
```

