---
title: "CA"
author: "Dennis Jakob"
date: "1/15/2020"
output: html_document
---
```{r}
require(vegan)
require(corrgram)
require(viridis)
require(scatterplot3d)
require(tidyverse)
data("varespec")
```

# CA exercise

1) What is the difference between the CA and PCA?

- PCA can only be used for data with linear relations between the variables
- CA can be used for unimodal data (non-linear) by using Chi^2 statistics

2) Perform a CA based on the varespec dataset
```{r}
# perform hellinger transformnation
varespec.hel <- decostand(varespec, "hellinger")
matplot(varespec,
        type = "l",
        pch=1,
        col = 1:dim(varespec)[2])
matplot(varespec.hel,
        type = "l",
        pch=1,
        col = 1:dim(varespec)[2])

# perform CA
varespec.cca <- cca(varespec)
summary(varespec.cca)
varespec.cca$CA$tot.chi
(sum(varespec.cca$CA$eig[1:2])/sum(varespec.cca$CA$eig))*100

varespec.hel.cca <- cca(varespec.hel)
summary(varespec.hel.cca)
varespec.hel.cca$CA$tot.chi
(sum(varespec.hel.cca$CA$eig[1:2])/sum(varespec.hel.cca$CA$eig))*100
```
helinger transformation is often used for bacterial abundances. removes also zeros.

3) Plot the first two CA axes
3.1) How much inertia is captured in these 2 axes?
3.2) What is the total inertia?
```{r}
plot(varespec.hel.cca,
     choices = c(1,2),
     scaling = 2)
(sum(varespec.hel.cca$CA$eig[1:2])/sum(varespec.hel.cca$CA$eig))*100
sum(varespec.hel.cca$CA$eig)*100
```

4) Apply a chi-squared test on the varespec dataset.
```{r}
chisq.test(varespec)
```
