---
title: "Ex_dbRDA"
author: "Dennis Jakob"
date: "1/15/2020"
output: html_document
---
# dbRDA exercise

1) What is the difference between RDA and db-RDA?
```{r}
require(vegan)
require(corrgram)
trees <- read.csv("statsII/trees.csv", header = T)

# prepare the data
row.names(trees) <- trees$ECOSYS
spec <- trees[,11:23]
env <- trees[,3:10]

# rows with only zeros cannot be handled by capscale
spec = spec + 0.001
```

2) Perform and plot an dbRDS using the trees.csv data set. Use the rankindex function to select an appropriate dissimilarity measures. Examine the data to determine which are the explanatory and response variables.
```{r}
# select a dissimilarity index
rankindex(grad = env,
          veg  = spec,
          method = "spearman")

# perform dbRDA
spec.dbRDA <- capscale(spec ~ .,
                       env,
                       dist = "bray") 
ordiplot(spec.dbRDA,
         type="t",
         scaling=1)
ordiplot(spec.dbRDA,
         type="t",
         scaling=2)
```

3) How much are the constrained and unconstrained variabilities?
```{r}
# check variabilities
spec.dbRDA$CCA$tot.chi
spec.dbRDA$CA$tot.chi
spec.dbRDA$tot.chi

# check constrained variance
spec.dbRDA$CCA$tot.chi/spec.dbRDA$tot.chi
# check unconstrained variance
spec.dbRDA$CA$tot.chi/spec.dbRDA$tot.chi
```

4) Test the significance of the model and the individual variables.
```{r}
# permutation tests to access significance of constraints
anova(spec.dbRDA) # overall test of the significance 
anova(spec.dbRDA,
      by="terms",
      perm.max=999)
```

5) Compute the correlation between the dbRDA first two axes and the explanatory variables.
```{r}
# correlation of the variables to the first two dbRDA axes
spec.dbRDA.species.scores = scores(spec.dbRDA,
                                   choices = c(1:2),
                                   display = "sites")
cor(cbind(spec.dbRDA.species.scores,
          env))
corrgram(cbind(spec.dbRDA.species.scores,
          env))
```

