---
title: "NMDS"
author: "Dennis Jakob"
date: "1/14/2020"
output: html_document
---
```{r}
require(vegan)
require(corrgram)
require(viridis)
require(scatterplot3d)
require(tidyverse)
data("varespec")
```

# NMDS exercise
1) Perform an NMDS analysis using the varespec dataset, based on the Bray-Curtis distance (see ?metaMDS).
```{r}
# compute distance matrix based on Bray Curtis distances
varespec.bc.dist.nmds2 <- metaMDS(comm = varespec,
                                 distance ="bray",
                                 k = 2,
                                 trymax = 1000,
                                 trace = F)
varespec.bc.dist.nmds2$stress
```


2) Plot the results. What can you say about the different sampling sites?
```{r}
plot(varespec.bc.dist.nmds,
     type = "t")
```

- not much can be said about sampling site as NMDS does not resolve any grouping

3) Did the algorithm converge?
```{r}
varespec.bc.dist.nmds$converged
```

4) Construct four NMDS varying from 1 to 4 the number of dimensions allowed, and make a plot of the stress values vs the number of dimensions.
```{r}
varespec.bc.dist.nmds1 <- metaMDS(comm = varespec,
                                 distance ="bray",
                                 k = 1,
                                 trymax = 1000,
                                 trace = F)
varespec.bc.dist.nmds3 <- metaMDS(comm = varespec,
                                 distance ="bray",
                                 k = 3,
                                 trymax = 1000,
                                 trace = F)
varespec.bc.dist.nmds4 <- metaMDS(comm = varespec,
                                 distance ="bray",
                                 k = 4,
                                 trymax = 1000,
                                 trace = F)
varespec.bc.dist.nmds1$stress
varespec.bc.dist.nmds3$stress
varespec.bc.dist.nmds4$stress

plot(seq(1,4,1),
     c(varespec.bc.dist.nmds1$stress,
       varespec.bc.dist.nmds2$stress,
       varespec.bc.dist.nmds3$stress,
       varespec.bc.dist.nmds4$stress))
```


5) Perform a Shepard plot. (see ?stressplot).
```{r}
stressplot(varespec.bc.dist.nmds1)
stressplot(varespec.bc.dist.nmds2)
stressplot(varespec.bc.dist.nmds3)
stressplot(varespec.bc.dist.nmds4)
```

```{r}
varespec.bc.dist.nmds10 <- metaMDS(comm = varespec,
                                    distance ="bray",
                                    k = 10,
                                    trymax = 100,
                                    trace = F)
scatterplot3d(x = varespec.bc.dist.nmds3$points[,1], y  = varespec.bc.dist.nmds3$points[,2], z = varespec.bc.dist.nmds3$points[,3])
plot(varespec.bc.dist.nmds10,
     type = "t")
stressplot(varespec.bc.dist.nmds10)

p <- plot_ly(x = varespec.bc.dist.nmds3$points[,1], y = varespec.bc.dist.nmds3$points[,2], z = varespec.bc.dist.nmds3$points[,3])
```

