---
title: "PCoA"
author: "Dennis Jakob"
date: "1/15/2020"
output: html_document
---
```{r}
require(vegan)
require(corrgram)
require(viridis)
require(scatterplot3d)
require(tidyverse)
data("varespec")
```

# PCoA exercise

1) Compute a distance matrix for the varespec dataset based on the Bray-Curtis dissimilarity measure.
```{r}
# compute distance matrix based on Bray Curtis distances
varespec.bc.dist <- vegdist(varespec,
                            method = "bray")
```

2) Perform a PCoA on the distance matrix and the plot the first two axes.
```{r}
varespec.bc.dist.PCoA <- cmdscale(varespec.bc.dist,
                                  k = 2,
                                  eig = T)
summary(varespec.bc.dist.PCoA)

x <- varespec.bc.dist.PCoA$points[,1]
y <- varespec.bc.dist.PCoA$points[,2]
plot(x, y, pch = 19, xlim=range(x) + c(-0.1,0.1), ylim=range(y)+c(-0.1,0.1))
text(x, y, pos = 4, labels = rownames(varespec), cex = 0.8)
```

3) Inspect the eigenvalues of the PCoA axes. Is there something unexpected?
4) Make an appropriate transformation of the distance matrix, apply PCoA and repeat tasks 1 to 3.
```{r}
# 3) negative values
barplot(varespec.bc.dist.PCoA$eig)

# 4)
varespec.bc.dist.cor.PCoA <- cmdscale(varespec.bc.dist,
                                  k = 2,
                                  eig = T,
                                  add = T)
summary(varespec.bc.dist.cor.PCoA)

x <- varespec.bc.dist.cor.PCoA$points[,1]
y <- varespec.bc.dist.cor.PCoA$points[,2]
plot(x, y, pch = 19, xlim=range(x) + c(-0.1,0.1), ylim=range(y)+c(-0.1,0.1))
text(x, y, pos = 4, labels = rownames(varespec), cex = 0.8)
barplot(varespec.bc.dist.cor.PCoA$eig)
# also a sqrt transformation works to turn eigenvalues positive
```

5) What happens if the distance matrix is based on Euclidean distance?
```{r}
varespec.euc.dist <- vegdist(varespec,
                            method = "euclidean")
varespec.euc.dist.PCoA <- cmdscale(varespec.euc.dist,
                                  k = 2,
                                  eig = T)
summary(varespec.euc.dist.PCoA)

barplot(varespec.euc.dist.PCoA$eig)

x <- varespec.euc.dist.PCoA$points[,1]
y <- varespec.euc.dist.PCoA$points[,2]
plot(x, y, pch = 19, xlim=range(x) + c(-0.1,0.1), ylim=range(y)+c(-0.1,0.1))
text(x, y, pos = 4, labels = rownames(varespec), cex = 0.8)
```

- eigenvalues do not become negative
- tail is a lot longer and flatter because the zeros are taken into account
- better separation of sites with low lycan coverage