---
title: "culturability"
author: "Dennis Jakob"
date: "11/12/2019"
output: html_document
---

# Determination of culturability of microbes from sylt seawater samples

## Experimenal setup on Sylt

Seawater was samples on different sites during the Sylt excursion 2019.

1. water sample from seawater pipeline of the AWI
2. water sample from the yard slip of List (Sylt)
3. water sample taken during the cruise with MyaII

The water samples were transfered in dilutions to Sy19 agar plates with a 96-pin holder to determine CFU.
All water samples were diluted as followed:

- 1:2
- 1:4
- 1:8
- 1:16

Estimated volume on plates: ~ 96 $\mu l$
The plates were incubation at RT for 3-4 weeks.

## Data analysis

```{r import cell count data}
cell_counts <- read.delim("cell_counts.txt", header = T, sep = "\t", dec = ",")
cell_counts <- cell_counts[,2:19]
```

```{r yard slip - dilution 1:16}
yardslip_16 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 16 & cell_counts[,"sample.site"] == "yard slip")

for (i in 3:ncol(yardslip_16)) {
  yardslip_16[5,i] <- mean(yardslip_16[1:4,i])
}
yardslip_16[6,3:18] <- (yardslip_16[5,3:18])/96
yardslip_16[7,3:18] <- (yardslip_16[6,3:18])*1000
yardslip_16[8,3:18] <- (yardslip_16[7,3:18])*16

yardslip_16[5:8,"sample.site"] <- yardslip_16[1,"sample.site"]
yardslip_16[5:8,"dilution.factor"] <- yardslip_16[1,"dilution.factor"]
rownames(yardslip_16) <- c("rep1", "rep2", "rep3", "rep4", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
yardslip_16[,3:18] <- round(x = yardslip_16[,3:18], digits = 0)
```

```{r yard slip - dilution 1:8}
yardslip_8 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 8 & cell_counts[,"sample.site"] == "yard slip")

for (i in 3:ncol(yardslip_8)) {
  yardslip_8[5,i] <- mean(yardslip_8[1:4,i])
}
yardslip_8[6,3:18] <- (yardslip_8[5,3:18])/96
yardslip_8[7,3:18] <- (yardslip_8[6,3:18])*1000
yardslip_8[8,3:18] <- (yardslip_8[7,3:18])*8

yardslip_8[5:8,"sample.site"] <- yardslip_8[1,"sample.site"]
yardslip_8[5:8,"dilution.factor"] <- yardslip_8[1,"dilution.factor"]
rownames(yardslip_8) <- c("rep1", "rep2", "rep3", "rep4", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
yardslip_8[,3:18] <- round(x = yardslip_8[,3:18], digits = 0)
```

```{r yard slip - dilution 1:4}
yardslip_4 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 4 & cell_counts[,"sample.site"] == "yard slip")

for (i in 3:ncol(yardslip_4)) {
  yardslip_4[5,i] <- mean(yardslip_4[1:4,i])
}
yardslip_4[6,3:18] <- (yardslip_4[5,3:18])/96
yardslip_4[7,3:18] <- (yardslip_4[6,3:18])*1000
yardslip_4[8,3:18] <- (yardslip_4[7,3:18])*4

yardslip_4[5:8,"sample.site"] <- yardslip_4[1,"sample.site"]
yardslip_4[5:8,"dilution.factor"] <- yardslip_4[1,"dilution.factor"]
rownames(yardslip_4) <- c("rep1", "rep2", "rep3", "rep4", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
yardslip_4[,3:18] <- round(x = yardslip_4[,3:18], digits = 0)
```

```{r yard slip - dilution 1:2}
yardslip_2 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 2 & cell_counts[,"sample.site"] == "yard slip")

for (i in 3:ncol(yardslip_2)) {
  yardslip_2[5,i] <- mean(yardslip_2[1:4,i])
}
yardslip_2[6,3:18] <- (yardslip_2[5,3:18])/96
yardslip_2[7,3:18] <- (yardslip_2[6,3:18])*1000
yardslip_2[8,3:18] <- (yardslip_2[7,3:18])*2

yardslip_2[5:8,"sample.site"] <- yardslip_2[1,"sample.site"]
yardslip_2[5:8,"dilution.factor"] <- yardslip_2[1,"dilution.factor"]
rownames(yardslip_2) <- c("rep1", "rep2", "rep3", "rep4", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
yardslip_2[,3:18] <- round(x = yardslip_2[,3:18], digits = 1)
```

```{r}
yardslip_cfu <- rbind(yardslip_16["cfu/ml seawater",], yardslip_8["cfu/ml seawater",], yardslip_4["cfu/ml seawater",], yardslip_2["cfu/ml seawater",])
```

```{r tab - dilution 1:16}
tab_16 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 16 & cell_counts[,"sample.site"] == "tab")

for (i in 3:ncol(tab_16)) {
  tab_16[5,i] <- mean(tab_16[1:4,i])
}
tab_16[6,3:18] <- (tab_16[5,3:18])/96
tab_16[7,3:18] <- (tab_16[6,3:18])*1000
tab_16[8,3:18] <- (tab_16[7,3:18])*16

tab_16[5:8,"sample.site"] <- tab_16[1,"sample.site"]
tab_16[5:8,"dilution.factor"] <- tab_16[1,"dilution.factor"]
rownames(tab_16) <- c("rep1", "rep2", "rep3", "rep4", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
tab_16[,3:18] <- round(x = tab_16[,3:18], digits = 0)
```

```{r tab - dilution 1:8}
tab_8 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 8 & cell_counts[,"sample.site"] == "tab")

for (i in 3:ncol(tab_8)) {
  tab_8[5,i] <- mean(tab_8[1:4,i])
}
tab_8[6,3:18] <- (tab_8[5,3:18])/96
tab_8[7,3:18] <- (tab_8[6,3:18])*1000
tab_8[8,3:18] <- (tab_8[7,3:18])*8

tab_8[5:8,"sample.site"] <- tab_8[1,"sample.site"]
tab_8[5:8,"dilution.factor"] <- tab_8[1,"dilution.factor"]
rownames(tab_8) <- c("rep1", "rep2", "rep3", "rep4", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
tab_8[,3:18] <- round(x = tab_8[,3:18], digits = 0)
```

```{r tab - dilution 1:4}
tab_4 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 4 & cell_counts[,"sample.site"] == "tab")

for (i in 3:ncol(tab_4)) {
  tab_4[5,i] <- mean(tab_4[1:4,i])
}
tab_4[6,3:18] <- (tab_4[5,3:18])/96
tab_4[7,3:18] <- (tab_4[6,3:18])*1000
tab_4[8,3:18] <- (tab_4[7,3:18])*4

tab_4[5:8,"sample.site"] <- tab_4[1,"sample.site"]
tab_4[5:8,"dilution.factor"] <- tab_4[1,"dilution.factor"]
rownames(tab_4) <- c("rep1", "rep2", "rep3", "rep4", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
tab_4[,3:18] <- round(x = tab_4[,3:18], digits = 0)
```

```{r tab - dilution 1:2}
tab_2 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 2 & cell_counts[,"sample.site"] == "tab")

for (i in 3:ncol(tab_2)) {
  tab_2[5,i] <- mean(tab_2[1:4,i])
}
tab_2[6,3:18] <- (tab_2[5,3:18])/96
tab_2[7,3:18] <- (tab_2[6,3:18])*1000
tab_2[8,3:18] <- (tab_2[7,3:18])*2

tab_2[5:8,"sample.site"] <- tab_2[1,"sample.site"]
tab_2[5:8,"dilution.factor"] <- tab_2[1,"dilution.factor"]
rownames(tab_2) <- c("rep1", "rep2", "rep3", "rep4", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
tab_2[,3:18] <- round(x = tab_2[,3:18], digits = 1)
```

```{r}
tab_cfu <- rbind(tab_16["cfu/ml seawater",], tab_8["cfu/ml seawater",], tab_4["cfu/ml seawater",], tab_2["cfu/ml seawater",])
```

```{r boat - dilution 1:16}
boat_16 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 16 & cell_counts[,"sample.site"] == "boat")

for (i in 3:ncol(boat_16)) {
  boat_16[4,i] <- mean(boat_16[1:3,i])
}
boat_16[5,3:18] <- (boat_16[4,3:18])/96
boat_16[6,3:18] <- (boat_16[5,3:18])*1000
boat_16[7,3:18] <- (boat_16[6,3:18])*16

boat_16[4:7,"sample.site"] <- boat_16[1,"sample.site"]
boat_16[4:7,"dilution.factor"] <- boat_16[1,"dilution.factor"]
rownames(boat_16) <- c("rep1", "rep2", "rep3", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
boat_16[,3:18] <- round(x = boat_16[,3:18], digits = 0)
```

```{r boat - dilution 1:8}
boat_8 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 8 & cell_counts[,"sample.site"] == "boat")

for (i in 3:ncol(boat_8)) {
  boat_8[4,i] <- mean(boat_8[1:3,i])
}
boat_8[5,3:18] <- (boat_8[4,3:18])/96
boat_8[6,3:18] <- (boat_8[5,3:18])*1000
boat_8[7,3:18] <- (boat_8[6,3:18])*8

boat_8[4:7,"sample.site"] <- boat_8[1,"sample.site"]
boat_8[4:7,"dilution.factor"] <- boat_8[1,"dilution.factor"]
rownames(boat_8) <- c("rep1", "rep2", "rep3", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
boat_8[,3:18] <- round(x = boat_8[,3:18], digits = 0)
```

```{r boat - dilution 1:4}
boat_4 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 4 & cell_counts[,"sample.site"] == "boat")

for (i in 3:ncol(boat_4)) {
  boat_4[4,i] <- mean(boat_4[1:3,i])
}
boat_4[5,3:18] <- (boat_4[4,3:18])/96
boat_4[6,3:18] <- (boat_4[5,3:18])*1000
boat_4[7,3:18] <- (boat_4[6,3:18])*4

boat_4[4:7,"sample.site"] <- boat_4[1,"sample.site"]
boat_4[4:7,"dilution.factor"] <- boat_4[1,"dilution.factor"]
rownames(boat_4) <- c("rep1", "rep2", "rep3", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
boat_4[,3:18] <- round(x = boat_4[,3:18], digits = 0)
```

```{r boat - dilution 1:2}
boat_2 <- subset(cell_counts, cell_counts[,"dilution.factor"] == 2 & cell_counts[,"sample.site"] == "boat")

for (i in 3:ncol(boat_2)) {
  boat_2[4,i] <- mean(boat_2[1:3,i])
}
boat_2[5,3:18] <- (boat_2[4,3:18])/96
boat_2[6,3:18] <- (boat_2[5,3:18])*1000
boat_2[7,3:18] <- (boat_2[6,3:18])*2

boat_2[4:7,"sample.site"] <- boat_2[1,"sample.site"]
boat_2[4:7,"dilution.factor"] <- boat_2[1,"dilution.factor"]
rownames(boat_2) <- c("rep1", "rep2", "rep3", "mean", "cfu/ul", "cfu/ml sample", "cfu/ml seawater")
boat_2[,3:18] <- round(x = boat_2[,3:18], digits = 1)
```

```{r}
boat_cfu <- rbind(boat_16["cfu/ml seawater",], boat_8["cfu/ml seawater",], boat_4["cfu/ml seawater",], boat_2["cfu/ml seawater",])
```

```{r}
cfu_total_rel <- rbind(yardslip_cfu, tab_cfu, boat_cfu)
rownames(cfu_total_rel) <- seq(1,12,1)

cfu_total_abs <- cbind(cfu_total_rel[,1:2],cfu_total_rel[,3:18]*100/10^6)


write.table(cfu_total_rel, "culturability_rel.txt", sep = "\t")
write.table(cfu_total_abs, "culturability_abs.txt", sep = "\t")
```

